Utilizing Parallel Workers: \\LLVM's Vectorization Plan

### 2024-11-19 paper/paper.tex
paper

\section{Introduction}
\label{sec:introduction}
Lorem ipsum

>>>>

\section{Introduction}
\label{sec:introduction}
Modern CPUs are often equipped with multiple different vector registers. These registers can nowadays 
be as wide as 512 bit, allowing for the processing of multiple data streams at once (SIMD). By 
batching multiple values together in one register, different ISAs like Intel AVX or ARM SVE allow 
the execution of the same instruction for all values in a vector register at once. Utilizing this 
can often lead to significant performance improvements over the scalar equivalent.

However, as manual code vectorization can quickly become very time consuming especially when
supporting different CPU architectures, modern compilers aim to automatically transform scalar code
to use vectorization when applicable.

As one of the most widely used compilation frameworks LLVM~\cite{10.5555/977395.977673} had 
implemented and refined its auto-vectorization over many years. It provides two different 
Vectorizers, one for innermost loops (LoopVectorize) and one for super-word parallelism 
(SLPVectorize)~\cite{llvmvec}.

This system however had quite a few limitations, as the loop vectorizer could only handle
innermost loops and neither outer loops, complex control flow, or non-inlined functions. 
Additionally, while multiple different vectorizatons for the same scalar code 
would be possible, the current vectorizers working directly on the LLVM IR had no capability 
of modelling and comparing the costs of such different vectorization approaches.

With these limitations in mind Intel started an ongoing refactorization effort to migrate LLVM's
auto-vectorization pipeline to use a more abstract Vectorization Plan 
(VPlan)~\cite{llvmextloopvec,llvmvplan}. The final goal would be to unite LLVM's auto-vectorization
in a single flexible system capable of optimizing SLPs, inner and nested loops with complex 
control flows. The auto-vectorization pass would create, compare, and transform multiple different 
VPlans each modelling a different vectorization approach and finally materialize the best one into
the LLVM IR.

In its current state LLVM's VPlan system is being integrated into the existing Loop Vectorizer and
already models most of the inner loop vectorization and transformations. Vectorization for outer 
loops is also in development and can be enabled by setting the \texttt{-enable-vplan-native-path} 
flag ~\cite{llvmouterloop}. In the future the plan will be to merge both of these loop vectorization
paths into one.
